---
title: "Cluster CEs related data"
output: html_notebook
---

# Init
```{r}
library(tidyverse)
library(magrittr)
library(readr)
library(tibble)
library(factoextra)
```

# Load data
Loads the data from the csv files.
```{r}
original_data <- read_csv("../data/nicepr_obs.csv")

ce_data <- read_csv("../data/nicepr_eval.csv")
```

# Data mumbojumbo
Transform original data to long format and add a type column to differentiate 
between original and counterfactual data.
```{r}
original_data %<>%
  pivot_longer(cols = starts_with("week"), 
               names_to = "week",
               values_to = "value") %>%
  mutate(week = str_replace(week, "minus", "-"),
         week = as.numeric(str_remove(week, "week_"))
        ) %>%
   mutate(type = "original")

ce_data %<>%
  pivot_longer(cols = starts_with("week"),
               names_to = "week",
               values_to = "value") %>%
  mutate(week = str_replace(week, "minus", "-"),
         week = as.numeric(str_remove(week, "week_"))
        ) %>%
  mutate(type = "ce") %>%
  select(id, week, value, type)

final_data <-
    original_data %>%
    bind_rows(ce_data)
```

# Clustering data
Transform the data into the wide format, compute the features (time series)
and normalize them.
```{r}
# compute the sum of difference clicks by periods between tma_weeks
normalize <- function(x){
  return((x - min(x))/(max(x) - min(x) + .Machine$double.xmin))
}

cluster_data <-
    final_data %>% 
    pivot_wider(names_from = type, 
                values_from = value,
                values_fn = {mean},
                values_fill = list(values = 0)) %>%
    mutate(diff = ce - original,
           ratio = ce/(original + .Machine$double.xmin),
           ratio = if_else(abs(ratio) == Inf, 200, ratio)
           ) %>% 
    pivot_wider(names_from = week, 
                names_glue = "week_{str_replace(as.character(week),'-','minus_')}_{.value}",
                values_from = c(original, ce, diff, ratio),
                values_fn = {mean},
                values_fill = list(values = 0)) %>% 
  rename(id_student = id) %>% 
  mutate(across(-id_student, ~ normalize(.)))
```

# Clustering
At first select number of clusters using the elbow method and silhouette method.
Then cluster the data using kmeans with k=2. Finally, visualize the clusters and 
cluster centers.

> NOTE: cluster_data_3 and strange naming conventions are results of the initial experiments

```{r}
cluster_data3 <-
  cluster_data %>%
  select(id_student, contains("diff"), contains("ratio"))

g_wss <- fviz_nbclust(cluster_data3[,-1],kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2, size=1.5) + 
  theme(plot.title = element_blank(),text = element_text(size = 12)) +
  ylab("WCSS")

g_sil <- fviz_nbclust(cluster_data3[,-1], kmeans, method = "silhouette") +
  theme(plot.title = element_blank(),text = element_text(size = 12))

g_measures <- g_wss / g_sil

ggsave("../output/cluster_measures.png", g_measures, width = 10, height = 5, dpi = 300)

set.seed(2206)
kmeans_2_2 <-
  cluster_data3[,-1] %>% 
  kmeans(2)

g_clusters <- fviz_cluster(kmeans_2_2, cluster_data3[,-1], stand = FALSE,
                           geom = "point", main = NULL) +
  theme_bw() +
  xlab("Dimension 1 (90.5%)") +
  ylab("Dimension 2 (3.7%)") +
  theme(text = element_text(size = 12),
        legend.text = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        legend.title = element_blank())

ggsave("../output/clusters.png", g_clusters, width = 10, height = 10, dpi = 300)

g_centers <- 
  kmeans_2_2$centers %>% 
  as_tibble() %>%
  mutate(cluster = paste0("cluster ",row_number())) %>% 
  pivot_longer(cols = -cluster, names_to = "variable", values_to = "value") %>% 
  mutate(variable = str_replace(variable, "minus_","-"),
         variable = str_replace(variable, "week_",""),
         week = purrr::map_dbl(variable,~as.numeric(str_split(., "_",simplify = TRUE)[1,1])),
         var = purrr::map_chr(variable, ~str_split(., "_",simplify = TRUE)[1,2])) %>% 
  select(cluster, week, var, value) %>%
  arrange(cluster, week, var) %>%
  mutate(Cluster = as.factor(cluster),
         var = if_else(var == "diff", "Difference in counterfactual and original activity counts", "Ratio between counterfactual and original activity counts")) %>% 
  ggplot(aes(x = week, y = value, color = Cluster)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~var, ncol = 1) + 
  scale_x_continuous(breaks = seq(-5, 40, by = 5),
                     expand = c(0,0)) +
  theme_bw() +
  xlab("Week") +
  ylab("Normalized feature value") +
  theme(text = element_text(size = 12),
        legend.text = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        legend.position = "bottom",
        legend.title = element_blank())

ggsave("../output/cluster_centers.png",g_centers, width = 10, height = 5, dpi = 300)
```